---
title: "MyDeepHealth"
output: html_notebook
---
Let's start with the acquisition of the 'Diabetes-Training' file first.

```{r}
df <- read.csv(file.choose(), header = T)
head(df)
```

Here we are calling all the column names to see what dependent or independent variables we will be working with.
```{r}
names(df)
```

Data Cleaning: we will analyze each variable for missing values using the formula below.
```{r}
any(is.na(df$race))
any(is.na(df$gender))
any(is.na(df$age))
any(is.na(df$weight))
```

since the 'Admission_type_id' variable has three missing values, we will replace each missing value with the mode (by creating a function first) since we are dealing with a categorical variable.
```{r}
any(is.na(df$admission_type_id))
```

Let's create a function that will calculate the mode and replace all missing values by that mode.
```{r}
getmode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}
df[is.na(df[,"admission_source_id"]), "admission_source_id"] <- getmode(df[,"admission_source_id"])
subset(df, is.na(df$admission_source_id))["admission_source_id"]
```

Let's move onto the next variable 'time_in_hospital'. This is a numerical variable, which means that its three missing values will be replaced by the variable's particular mean.
```{r}
any(is.na(df$time_in_hospital))
sum(is.na(df$time_in_hospital))
meantime <- mean(df$time_in_hospital, na.rm = T)
getmean <- round(meantime,0)
getmean
df[is.na(df[,"time_in_hospital"]), "time_in_hospital"] <- mean(df$time_in_hospital, na.rm = T) + getmean
```
Since we have determined the mean and rounded it to one sigfig, we will replace the missing values with the mean here below.
```{r}
subset(df, is.na(df$time_in_hospital))["time_in_hospital"]
```
To make things easy for ourselves, let's identify both numerical (or continuous) and categorical variables and create two vectors comprising one of each tyoe of data structure.
```{r}
categorical_col <- c("discharge_disposition_id", "admission_type_id", "medical_specialty", "num_procedures", 
                     "diag_1", "diag_2","diag_3", "max_glu_serum", "A1Cresult", "metformin", "repaglinide", 
                     "nateglinide", "chlorpropamide", "glimepiride", "acetohexamide", "glipizide", "glyburide", 
                     "tolbutamide", "pioglitazone", "rosiglitazone", "acarbose", "miglitol", "troglitazone", 
                     "tolazamide", "examide", "citoglipton", "insulin", "glyburide.metformin", "glipizide.metformin", 
                     "glimepiride.pioglitazone", "metformin.rosiglitazone", "metformin.pioglitazone", "change", 
                     "diabetesMed", "readmitted")
numerical_col <- c("num_lab_procedures", "num_medications", "number_outpatient", "number_emergency", "number_inpatient", 
                   "number_diagnoses")
```

Since categorical and numerical variables are now separated in two different vectors, let's see if we can determine if a certain variable has missing values, and if so, replace those values with either the mode or mean.
First with the categorical vector, we will build a function that will determine missingness and replace it with mode.
Now will identify all NA for numerical variables and take care of them.
```{r}
cat_fun <- function(variable,df) {
  n <- sum(is.na(df[variable])) 
  if (n > 0) {
    m <- getmode(df[,variable])
    df[is.na(df[,variable]), variable] <- m
  } 
  df[variable]
}
for(x in categorical_col) {
  result <- cat_fun(x,df)
  print(result)
} 
```
```{r}
subset(df, is.na(df$medical_specialty))["medical_specialty"]
```

We can use the above line to check if every missing values in categorical variables has been taken care of. 
We have successfully replaced every missing value with the mode in our categorical variables.
Let's take care of missing values for continuous variables as well.
```{r}
cont.fun <- function(z,df) {
  s <- sum(is.na(df[z]))
  if (s > 0) {
    t <- mean(df[,z])
    df[is.na(df[,z]), z] <- t
  }
  df[z]
}
for(z in numerical_col) {
  feedback <- cont.fun(z,df)
  print(feedback)
}
```

```{r}
subset(df, is.na(df$number_outpatient))["number_outpatient"]
```
Let's ditermine the probability distribution for readmission by first determining the frequencies of each of the three entries in the re-admitted column.
```{r}
e <- table(df$readmitted)
d <- length(df$readmitted)
e/d
```
Feature Engineering-- to be done 
A. Exploratory data analysis 
Let's determine correlation by first determining numerical colums in our dataset.
```{r}
require(dplyr)
```

```{r}
num.cols <- sapply(df,is.numeric)
num.cols
```
B. Correlation 
```{r}
cor.data <- cor(df[,num.cols])
print(cor.data)
```
Using Corrplot, we will be able to visualize correlation more clearly.
```{r}
require(corrplot)
print(corrplot(cor.data, method = 'color'))
```
Using a histogram, we will visualize the distribution of inputs in the target (dependent) variable.
As it turns out, our target variable is discrete therefore we cannot deduce a reliable set of information from it (this is a subject to be explored more in depth)
```{r}
require(ggplot2)
ggplot(df, aes(readmitted)) + geom_histogram(alpha = 0.5, fill = 'green')
```

